%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%	LaTeX File for Stanford University PhD Thesis
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\prefacesection{Acknowledgements}


I am grateful for everyone and everything that has contributed to the success of this thesis.  I am indebted to the supervision of Prof. Sanjay Chawla,   co-author Mr. Aditya Menon  for interesting research ideas and  guidance. I also appreciate Dr. Ehsan Zare Borzeshi  for providing constructive feedback for my thesis and I  thank Dr. Fabio Ramos for keeping me on track. Other mentions include my mentor at Capital Markets CRC Dr. Uma Srinivasan for very helpful PhD advice.


 I want to acknowledge the many opportunities from various organisations during my PhD;
  Capital Market Cooperative Research Centre (CMCRC) for funding the majority of my PhD,   NSW Ministry of Health, and NSW Ambulance for opportunity to learn practical skills as data science research officer, for allowing me to practise data science analysis skills as Data Scientist.  School of Information Technology for providing me the oppurtunity to visit, Qatar Computing Research Institute (QCRI)  for a short study trip in Qatar as well as Sanjay for providing me with the opportunity to present at the PAKDD conference  in Auckland.

Finally, I want to express my sincere gratitude for my family who have supported and stood beside's me along the way:
My mother Leelavathi, father Chalapathy - for being the source of inspiration to explore unknown frontier's.
My wife Mamatha - for putting up with my craziness and always being so encouraging and supportive;
My father in-law Seetharaman and mother in-law  Vasanthi- who believed in my abilities and have always encouraged me to move forward.
My sister's Vandana, Radhika who have silently backed me throughout. and lastly My lovely daughter Tushara, who with her smile has taught me, to be happy for no reason.

My PhD journey has taught me to never stop learning but also that we can find purpose even in the most difficult situations.
I am grateful to all my teachers and living beings who have enabled me to be where I am today.
